---
title: "Lab 5 Report"
author: "Amara Miller"
date: "`r format(Sys.Date())`"
output: github_document
---

## Loading Packages and Data

```{r}
library(tidyverse)
iris
iris <- as_tibble(iris)
iris
```


## Questions and Answers for Lab 5 Report

## Question 1

1. Rename each variable so that it is all lower-case and uses an underscore _ instead of a period . in the name (the recommended coding style in the tidyverse style guide). Print the resulting table.

```{r}
new_iris <- rename(iris, sepal_length = Sepal.Length, sepal_width = Sepal.Width, petal_length = Petal.Length, petal_width = Petal.Width, species = Species)
new_iris
```


## Question 2

2. Convert the four numerical variables from cm to mm by multiplying by 10. Print the resulting table.

```{r}
iris_cm <- mutate(new_iris, sepal_length = sepal_length * 10, sepal_width = sepal_width * 10, petal_length = petal_length * 10, petal_width = petal_width * 10)
iris_cm
```


## Question 3

3. Calculate sepal area and petal area (area is equal to length multiplied by width). Print a table with only the variables sepal area, petal area, and species.

```{r}
iris_area <-mutate(iris_cm, sepal_area = sepal_length * sepal_width, petal_area = petal_length * petal_width)
iris_onlyarea <- select(iris_area, sepal_area, petal_area, species)
iris_onlyarea
```

## Question 4

4. Calculate the following statistics for the entire dataset from the sepal length variable and print the resulting table:

sample size
maximum value
minimum value
range
median
first quartile (q1)
third quartile (q2)
inter-quartile range (iqr)

```{r}
summarise(
  iris,
  sample_size = n(),
  min = min(Sepal.Length),
  max = max(Sepal.Length),
  range = max(Sepal.Length) - min(Sepal.Length),
  median = median(Sepal.Length),
  q1 = quantile(Sepal.Length, probs = 0.25),
  q2 = quantile(Sepal.Length, probs = 0.75),
  IQR = IQR(Sepal.Length)
)
```


## Question 5

5. Calculate the following statistics for each species from the petal width variable and print the resulting table:

sample size
mean
standard deviation
variance
standard error of the mean
approximate 95% confidence interval

```{r}
iris_grouped <- group_by(iris, Species)
iris_grouped

Petal.Width_summary <-
  summarize(
    iris_grouped,
    sample_size = n(),
    mean_pw = mean(Petal.Width),
    sd = sd(Petal.Width),
    var = var(Petal.Width),
    sem = sd(Petal.Length) / sqrt(n()),
    ci_upper_limit = mean_pw + 1.96 * sem,
    ci_lower_limit = mean_pw - 1.96 * sem
  )
Petal.Width_summary

```


## Question 6 

6. Visualize the relationship between petal width and species using a strip plot.

```{r}
ggplot(data = iris) +
  geom_jitter(mapping = aes(x = Species, y = Petal.Width))
```


## Question 7

7. Starting with the previous graph, add the mean and 95% confidence interval for each species.

```{r}
ggplot(data = iris) +
  geom_jitter(mapping = aes(x = Species, y = Petal.Width)) +
  geom_crossbar(
    data = Petal.Width_summary, 
    mapping = aes(x = Species, y = mean_pw, ymax = ci_upper_limit, 
                  ymin = ci_lower_limit),
    color = "red"
  )
```


## Question 8 

8. Visualize the relationship between petal length, petal width, and species using a scatterplot. Map the two numerical variables to the x and y axes and map species to the color and shape aesthetics.
```{r}
ggplot(data = iris) +
  geom_point(mapping = aes(x = Petal.Length, y = Petal.Width, color = Species))
```


## Session Info

Here is my session information:

```{r}
sessioninfo::session_info()
```